---
title: "BDA_Assignment2"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r}
# Dataset on Amazon's Top 50 bestselling books from 2009 to 2019 contains 550 books. data has been categorized into fiction and non-fiction using Goodreads. The name of the dataset is bestsellers with categories.csv. No Null Values seen in the dataset. The data dictionary is seen as below.
# 
# Data Dictionary
# •	Name - name of the bestselling book (datatype - object)
# •	Author - author of the bestselling book (datatype - object)
# •	User Rating - User Rating of the book (datatype - float)
# •	Reviews - number of reviews for the book (datatype - integer)
# •	Price - price of the book (datatype - integer)
# •	Year - Year when it was a bestseller(datatype - integer)
# •	Genre - Categorised as Fiction and non fiction (datatype - object)
```

```{r setup, include=FALSE}
# First, import necessary packages and dataset.
library(flexdashboard)
library(dplyr)
library(ggplot2)
library(knitr)
library(DT)

# Read the dataset.
data <- read.csv("bestsellers with categories.csv")
```


Question1
=====================
Row
------------------------------

1. Create two graphs that show the most 10 popular authors and most 10 popular books. [30 Marks]
```{r}
# Create a graph for popular authors
popular_authors <- data %>%
  group_by(Author) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  head(10)

# Create a bar chart
plot1 <- ggplot(popular_authors, aes(x = reorder(Author, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "lightblue", color = "white") +
  labs(title = "Top 10 Popular Authors",
       x = "Author",
       y = "Number of Bestsellers") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(label = Count), vjust = -0.5) +
  theme_classic()
print(plot1)
```

Here I use count of books as the criteria of popular books.
```{r}
# Create a graph for popular books
popular_books <- data %>%
  group_by(Name) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  head(10)

# Create a bar chart
plot2 <- ggplot(popular_books, aes(x = Count, y = reorder(Name, -Count))) +
  geom_bar(stat = "identity", fill = "lightblue", color = "white") +
  labs(title = "Top 10 Popular Books",
       x = "Number of Bestsellers",
       y = "Book Name") +
  geom_text(aes(label = Count), hjust = -0.5) +
  theme_classic()

print(plot2)
```



Question2
=====================
Row
------------------------------

2. Create a table that shows all the best Books from 2009 to 2019 by users rating and reviews (rating ratings > 4.9 and number of reviews > 5,000). [20 Marks]
```{r}
best_books <- data %>%
  filter(User_Rating >= 4.9, Reviews > 5000, Year >= 2009, Year <= 2019) %>%
  arrange(desc(User_Rating))

DT::datatable(best_books,
              rownames = FALSE,
              options = list(pageLength = 25, scrollY = FALSE),
              class = 'display')
```



Question3
=====================
Row
------------------------------

3. Create a trendline that shows the numbers of reviews along with 2009 to 2019. [20 Marks]
```{r}
# Create a trendline for reviews
plot3 <- ggplot(data, aes(x = Year, y = Reviews)) +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Trend of Reviews Over Time",
       x = "Year",
       y = "Number of Reviews") +
  theme_light()
print(plot3)
```



Question4
=====================
Row
------------------------------

4. Create any table or graph of your interest. [10 Marks]

Pie Chart -- show the distribution of fiction and non-fiction books.
```{r}
fiction_nonfiction <- data %>%
  group_by(Genre) %>%
  summarise(Count = n())

pie_chart <- ggplot(fiction_nonfiction, aes(x = "", y = Count, fill = Genre)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(title = "Distribution of Fiction and Non-Fiction Books") +
  theme_linedraw() +
  scale_fill_manual(values = c('dodgerblue', "purple"))

print(pie_chart)
```



Question5
=====================
Row
------------------------------

5. Ensure the tables and graphs are suitable for effectively presenting the results, displaying a high level of organization and aesthetic appeal. [20 Marks]

(1) Boxplot -- display the distribution of book prices.
```{r}
boxplot_plot <- ggplot(data, aes(x = Genre, y = Price, fill = Genre)) +
  geom_boxplot() +
  labs(title = "Distribution of Book Prices by Genre",
       x = "Genre",
       y = "Price") +
  scale_fill_manual(values = c('skyblue', "lightgreen"))

print(boxplot_plot)
```

(2) Bar Chart -- display the number of fiction and non-fiction books by year.
```{r}
genre_by_year <- data %>%
  group_by(Year, Genre) %>%
  summarise(Count = n())

bar_chart <- ggplot(genre_by_year, aes(x = as.factor(Year), y = Count, fill = Genre)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Number of Fiction and Non-Fiction Books by Year",
       x = "Year",
       y = "Count") +
  scale_fill_manual(values = c('dodgerblue', "skyblue")) +
  theme_classic()

print(bar_chart)
```

(3) Scatter Plot -- visualize the relationship between book price and user rating.
```{r}
scatter_plot <- ggplot(data, aes(x = Price, y = User_Rating, color = Genre)) +
  geom_point() +
  labs(title = "Scatter Plot of Book Price vs. User Rating",
       x = "Book Price",
       y = "User Rating") +
  scale_color_manual(values = c("Fiction" = "dodgerblue", "Non Fiction" = "purple")) +
  theme_minimal()

print(scatter_plot)
```

