---
title: "BDA_Assignment2"
author: "Ruoxin WANG"
date: "2023-11-05"
output: html_document
---
Dataset on Amazon's Top 50 bestselling books from 2009 to 2019 contains 550 books. data has been categorized into fiction and non-fiction using Goodreads. The name of the dataset is bestsellers with categories.csv. No Null Values seen in the dataset. The data dictionary is seen as below.

Data Dictionary
•	Name - name of the bestselling book (datatype - object)
•	Author - author of the bestselling book (datatype - object)
•	User Rating - User Rating of the book (datatype - float)
•	Reviews - number of reviews for the book (datatype - integer)
•	Price - price of the book (datatype - integer)
•	Year - Year when it was a bestseller(datatype - integer)
•	Genre - Categorised as Fiction and non fiction (datatype - object)

First, import dataset.
```{r}
library(dplyr)
library(ggplot2)
library(knitr)

# Read the dataset
df <- read.csv("bestsellers with categories.csv")
```


Now you are a data analyst who wants to create a dashboard using R Flexdashboard for the dataset. The requirements include: 

1. Create two graphs that show the most 10 popular authors and most 10 popular books. [30 Marks]
```{r}
# Create a graph for popular authors
popular_authors <- df %>%
  group_by(Author) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  head(10)

# Create a bar chart
plot1 <- ggplot(popular_authors, aes(x = reorder(Author, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Top 10 Popular Authors",
       x = "Author",
       y = "Number of Bestsellers")
print(plot1)
```

```{r}
# Create a graph for popular books
popular_books <- df %>%
  group_by(Name) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  head(10)

# Create a bar chart
plot2 <- ggplot(popular_books, aes(x = reorder(Name, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "green") +
  labs(title = "Top 10 Popular Books",
       x = "Book Name",
       y = "Number of Bestsellers")
print(plot2)
```


2. Create a table that shows all the best Books from 2009 to 2019 by users rating and reviews (rating ratings > 4.9 and number of reviews > 5,000). [20 Marks]
```{r}
# # Filter the data to select best books based on rating and reviews criteria
# best_books <- df %>%
#   filter(User_Rating >= 4.9, Reviews > 5000, Year >= 2009, Year <= 2019) %>%
#   arrange(desc(User_Rating))
# 
# # Create a data frame
# best_books_df <- data.frame(
#   Name = best_books$Name,
#   Author = best_books$Author,
#   User_Rating = best_books$User_Rating,
#   Reviews = best_books$Reviews,
#   Year = best_books$Year
# )
# # Print the table in the command line
# print(best_books_df)

# # Subset the data to select best books based on rating and reviews criteria
# best_books <- subset(df, User_Rating >= 4.9 & Reviews > 5000 & Year >= 2009 & Year <= 2019)
# # Order the best books by user rating in descending order
# best_books <- best_books[order(-best_books$User_Rating), ]
# # Display the table of best books
# best_books


# Filter the data to select best books based on rating and reviews criteria
best_books <- df %>%
  filter(User_Rating >= 4.9, Reviews > 5000, Year >= 2009, Year <= 2019) %>%
  arrange(desc(User_Rating))

# Display the table of best books
best_books


# # Filter the data to select best books based on rating and reviews criteria
# best_books <- df[df$User_Rating >= 4.9 & df$Reviews > 5000 & df$Year >= 2009 & df$Year <= 2019, ]
# 
# # Order the best books by user rating in descending order
# #best_books <- best_books[order(-best_books$User_Rating), ]
# 
# # Display the table of best books
# best_books

```


3. Create a trendline that shows the numbers of reviews along with 2009 to 2019. [20 Marks]
```{r}
# Create a trendline for reviews
plot3 <- ggplot(df, aes(x = Year, y = Reviews)) +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Trend of Reviews Over Time",
       x = "Year",
       y = "Number of Reviews")
print(plot3)
```


4. Create any table or graph of your interest. [10 Marks]
Pie Chart -- show the distribution of fiction and non-fiction books.
```{r}
fiction_nonfiction <- df %>%
  group_by(Genre) %>%
  summarise(Count = n())

pie_chart <- ggplot(fiction_nonfiction, aes(x = "", y = Count, fill = Genre)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(title = "Distribution of Fiction and Non-Fiction Books")

print(pie_chart)
```

5. Ensure the tables and graphs are suitable for effectively presenting the results, displaying a high level of organization and aesthetic appeal. [20 Marks]
(1) Boxplot -- display the distribution of book prices.
```{r}
boxplot_plot <- ggplot(df, aes(x = Genre, y = Price, fill = Genre)) +
  geom_boxplot() +
  labs(title = "Distribution of Book Prices by Genre",
       x = "Genre",
       y = "Price")

print(boxplot_plot)
```

(2) Bar Chart -- display the number of fiction and non-fiction books by year.
```{r}
genre_by_year <- df %>%
  group_by(Year, Genre) %>%
  summarise(Count = n())

bar_chart <- ggplot(genre_by_year, aes(x = as.factor(Year), y = Count, fill = Genre)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Number of Fiction and Non-Fiction Books by Year",
       x = "Year",
       y = "Count")

print(bar_chart)
```

(3) Scatter Plot -- visualize the relationship between book price and user rating.
```{r}
scatter_plot <- ggplot(df, aes(x = Price, y = User_Rating, color = Genre)) +
  geom_point() +
  labs(title = "Book Price vs. User_Rating",
       x = "Price",
       y = "User_Rating")

print(scatter_plot)
```




