---
title: "Week 7 Materials"
output:
  html_document:
    df_print: paged
---
```{r}
library(tidyverse)
library(readxl)
library(lubridate)
options(scipen=100, digits=4)
```

```{r}
df <- data.frame(x=1:20, y=2*1:20+0.3*((1:20)^2) + rnorm(20, 0, 10))
df
```
#Aids the eye in seeing patterns in the presence of overplotting
#se： Display confidence interval around smooth?
#method: Smoothing method (function) to use, accepts either NULL or a character vector, e.g. "lm", "glm", "gam"、
```{r}
df %>%
  ggplot(aes(x,y)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) +
  theme_classic()
```

# formula: Formula to use in smoothing function, eg. y ~ x, y ~ poly(x, 2), y ~ log(x)
```{r}
df %>%
  ggplot(aes(x,y)) +
  geom_point() +
  geom_smooth(method="lm", se=TRUE, formula = y ~poly(x, 2)) +
  theme_classic()
```
#show the confidence level around smooth
```{r}
df %>%
  ggplot(aes(x,y)) +
  geom_point() +
  geom_smooth() +
  theme_classic()
```
# Example 8.3: Home Market Value Data
```{r}
op<-read_xlsx("Crude Oil Prices.xlsx", skip=3)
op<- op %>% mutate(Date=ym(Date))
op<-op %>% arrange(Date) %>% mutate(Time=row_number(Date))

```
# Exponential 
# Need to use nls; starting values for a & b are estimated by using OLS
```{r}
lmfit<-lm(log(Price)~Time,op)
A = coef(lmfit)[1]
B = coef(lmfit)[2]

ggplot_exp<-op %>%
  ggplot(aes(x=Time, y=Price)) +
  geom_line() +
  theme_classic()+
  geom_smooth(method='nls',formula=y~a*exp(b*x), se=FALSE,method.args = list(start = list(a = A, b = B)))
ggplot_exp
```
# Logarithmic 
```{r}
ggplot_log<-op %>%
  ggplot(aes(x=Time, y=Price)) +
  geom_line() +
  theme_classic()+
  geom_smooth(method="lm", formula=y~log(x), se=FALSE)
ggplot_log
```
# Polynomial 2nd  
```{r}
ggplot_poly2<-op %>%
  ggplot(aes(x=Time, y=Price)) +
  geom_line() +
  theme_classic()+
  geom_smooth(method='lm',formula=y~poly(x,2), se=FALSE)
ggplot_poly2
```
# Polynomial 3rd  
```{r}
ggplot_poly3<-op %>%
  ggplot(aes(x=Time, y=Price)) +
  geom_line() +
  theme_classic()+
  geom_smooth(method='lm',formula=y~poly(x,3), se=FALSE)
ggplot_poly3
```
# Power 
# similar to the exponential function case, nls is used for estimating a trendline model so starting values must be given to make the model converge.
```{r}
ggplot_power<-op %>%
  ggplot(aes(x=Time, y=Price)) +
  geom_line() +
  theme_classic()+
  geom_smooth(method='nls',formula=y~a*b^x, se=FALSE,,method.args=list(start=list(a=1, b=1)))
ggplot_power
```
# Present all graphs at once
```{r}
library(cowplot)
all_plots<-cowplot::plot_grid(ggplot_exp, ggplot_log, ggplot_poly2, ggplot_poly3, ggplot_power, labels = c("Exponential", "Logarithmic","Polynomial 2nd","Polynomial 3rd", "Power"), label_size=10, ncol=2)
all_plots
```
# Example 8.3: Home Market Value Data
```{r}
hv<-read_xlsx("Home Market Value.xlsx", skip=2)
```

```{r}
hv %>% ggplot(aes(`Square Feet`,`Market Value`))+
  geom_point()+
  theme_classic()
```
# In the case of simple regression, multiple R y~x = cor(x,y); r-squared y~x = cor(x,y)^2
```{r}
cor(hv$`Market Value`,hv$`Square Feet`)
cor(hv$`Market Value`,hv$`Square Feet`)^2
```
```{r}
reg<-lm(`Market Value`~`Square Feet`,hv)
summary(reg)
```
```{r}
fitted(reg)
plot(hv$`Square Feet`,hv$`Market Value`)
abline(reg)
```
```{r}
1-((42-1)/(42-1-1))*(1-0.5347)
```
```{r}
anova(reg)
```

# regression diagnostics
```{r}
par(mfrow=c(2,2))
plot(reg)
```

# Multiple regression 
```{r}
myData<-read_xlsx("Colleges and Universities.xlsx")
options(scipen=999)
Model <-lm(Earnings ~ Cost + Grad + Debt + City, data = myData)
summary(Model)

predict(Model, data.frame(Cost=25000, Grad=60, Debt=80, City=1), level=0.95, interval="confidence")

predict(Model, data.frame(Cost=25000, Grad=60, Debt=80, City=1), level=0.95, interval="prediction")
```



```{r}
cu<-read_xlsx("Colleges and Universities.xlsx", skip=2)
reg2<-lm(`Graduation %`~`Median SAT`+`Acceptance Rate`+`Expenditures/Student`+`Top 10% HS`,cu)
summary(reg2)
```
# Identifying the Best Regression Model (Exmple 8.13)
```{r}
ba<-read_xlsx("Banking Data.xlsx", skip=3)
```

```{r}
reg3<-lm(Balance~Age+Education+Income+`Home Value`+Wealth,ba)
summary(reg3)
```

```{r}
reg4<-lm(Balance~Age+Education+Income+Wealth, ba)
summary(reg4)
```

#------------------------------------Session 2----------------------

# Multicollinearity - an example    
```{r}
library(tidyverse)
library(readxl)
options(scipen=100, digits=4)
```
```{r}
cu<-read_xlsx("Colleges and Universities.xlsx", skip=2)
hv<-read_xlsx("Home Market Value.xlsx", skip=2)
```
```{r}
cor(cu %>% select(`Graduation %`,`Median SAT`,`Acceptance Rate`,`Expenditures/Student`,`Top 10% HS`))
```
```{r}
cor(ba %>% select(Balance,Age,Education,Income,Wealth))
```
# Example 8.14
```{r}
reg4<-lm(Balance~Age+Education+Income+Wealth, ba)
summary(reg4)
```
# w/o Wealth
```{r}
reg5<-lm(Balance~Age+Education+Income,ba)
summary(reg5)
```
# w/o Education
```{r}
reg6<-lm(Balance~Age+Income,ba)
summary(reg6)
```
# w/o Income
```{r}
reg7<-lm(Balance~Age+Education+Wealth,ba)
summary(reg7)
```

# A Model with Categorical Variables (Example 8.15)
```{r}
sd<-read_xlsx("Employee Salaries.xlsx", skip=2)
```
```{r}
reg8<-lm(Salary~Age+as.factor(MBA),sd)
summary(reg8)
```


# A Model with Categorical Variables with 2 + levels
```{r}
is<-read_xlsx("Insurance Survey.xlsx", skip=2, n_max=25)
```
```{r}
unique(is$Education)
```
```{r}
is <- is %>% mutate(graduate_degree=ifelse(Education=="Graduate degree",1,0))
is <- is %>% mutate(college_graduate=ifelse(Education=="College graduate",1,0))
is %>% select(Education, graduate_degree, college_graduate)
```

# A Model with Categorical Variables with 2+ levels - Example 8.17

```{r}
sf<-read_xlsx("Surface Finish.xlsx", skip=2)
```
```{r}
unique(sf$`Cutting Tool`)
```
```{r}
sf <- sf %>% mutate(TypeB=ifelse(`Cutting Tool`=="B",1,0))
sf <- sf %>% mutate(TypeC=ifelse(`Cutting Tool`=="C",1,0))
sf <- sf %>% mutate(TypeD=ifelse(`Cutting Tool`=="D",1,0))
sf %>% select(`Cutting Tool`, TypeB, TypeC, TypeD)
```
```{r}
reg9<-lm(`Surface Finish`~RPM+TypeB+TypeC+TypeD,sf)
summary(reg9)
```
```{r}
reg10<-lm(`Surface Finish`~RPM+as.factor(`Cutting Tool`),sf)
summary(reg10)
```
# Incorporating Interaction Terms in a Regression Model (Example 8.16)
```{r}
sd<-read_xlsx("Employee Salaries.xlsx", skip=2)
```
```{r}
sd <- sd %>% mutate(MBA_bi=ifelse(MBA=="Yes",1,0))
sd <- sd %>% mutate(Interaction=Age*MBA_bi)
reg11<-lm(Salary~Age+MBA_bi+Interaction,sd)
# reg11<-lm(Salary~Age+MBA_bi+Age*MBA_bi,sd)
summary(reg11)
```

```{r}
reg12<-lm(Salary~Age+Interaction,sd)
summary(reg12)
```

# Regression Models with Nonlinear Terms

```{r}
Year <- c(1959, 1960, 1961, 1962, 1963, 1964, 1965, 1966, 1967, 1968, 1969)
Population <- c(4835, 4970, 5085, 5160, 5310, 5260, 5235, 5255, 5235, 5210, 5175)
sample1 <- data.frame(Year, Population)
sample1$Year <- sample1$Year-1964
```

```{r}
ggplot(sample1, aes(Year, Population)) +
  geom_point() +
  theme_classic()
```
#ploynomial regression 
# The I() function treats the contents within the parentheses as an R expression.
```{r}
fit1 <- lm(Population ~ Year, data=sample1)
fit2 <- lm(Population ~ Year + I(Year^2), data=sample1)
fit3 <- lm(Population ~ Year + I(Year^2) + I(Year^3), data=sample1)
summary(fit1)
summary(fit2)
summary(fit3)

```
```{r}
ggplot(sample1, aes(Year, Population)) +
  geom_point() +
  geom_line(aes(y=predict(fit2))) +
  theme_classic()
```

# Modeling Beverage Sales Using Curvilinear Regression  (Example 8.18)
```{r}
bs<-read_xlsx("Beverage Sales.xlsx", skip=2)
```
```{r}
reg13<-lm(Sales~Temperature,bs)
summary(reg13)
```
```{r}
par(mfrow=c(2,2))
plot(reg13)
```
```{r}
reg14<-lm(Sales~Temperature+I(Temperature^2),bs)
# reg14<-lm(Sales~Temperature+Temperature*Temperature,bs)
summary(reg14)
```
```{r}
par(mfrow=c(2,2))
plot(reg14)
```