---
title: "Week 4 Material"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---



```{r}
rm(list = ls())
dt <- read_csv("projects.csv")
```

# proprotion is commonly used to deal with categorical data
```{r}
dt %>%
  group_by(poverty_level) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n))
```


```{r}
dt %>%
  select(id, grade_level, poverty_level)
```
# Contingency Table 列联表

```{r}
dt %>%
  select(grade_level, poverty_level) %>%
  table()
```

```{r}
dt %>%
  select(grade_level, poverty_level) %>%
  table() %>%
  prop.table()
```

```{r}
dt %>%
  select(poverty_level, resource_type, grade_level) %>%
  table()
```

```{r}
dt %>%
  select(grade_level, poverty_level) %>%
  filter(poverty_level != "unknown") 
```

# Pivot Table
```{r}
library(rpivotTable)
dt %>% 
  filter(poverty_level != "unknown")  %>%
  filter(is.na(grade_level)==FALSE)    %>%
  rpivotTable(rows=c("grade_level"), cols=c("poverty_level"))
```

# Measures of location: Application
```{r}
library(modeest)
```


# Measures of location: Application - Repair time
```{r}
library(readxl)
repair<-read_excel("Computer Repair Times.xlsx", skip=2,sheet=1)
repair <- repair %>% rename(time = `Repair Time (Days)`)
```


```{r}
repair %>% 
  ggplot(aes(x=time)) +
  geom_histogram(binwidth = 1) +
  theme_classic()
```
```{r}
repair %>%
  group_by(time) %>%
  summarise(count = n()) %>% 
  arrange(time) %>%
  mutate(cum_freq = cumsum(count)) %>%
  mutate(cum_freq = cumsum(count), rel_cum_freq = cum_freq/sum(count)) %>%
  ungroup()
```


```{r}
library(tidyverse)
library(lubridate) ## R package for managing date and time values
```

## if the above is not working:
```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(lubridate)
```



```{r}
dt <- read_csv("donors.csv")
```

```{r}
# 
dt %>% 
  select(state) %>%
  group_by(state) %>%
  summarise(n = n())
```

```{r}
dt %>% 
  select(state) %>%
  group_by(state) %>%
  summarise(n = n()) %>%
#generate a new column that equals to n/sum(n)
  mutate(freq = n / sum(n))
```

# Frequncy distribution for numerical (continous) values
```{r}
dt %>% 
  filter(num_projects <= 10) %>%
  ggplot(aes(x=num_projects)) +
  geom_histogram() +
  theme_classic()
```

# Histogram
```{r}
dt %>% 
  filter(num_projects <= 10) %>%
  group_by(group = cut(num_projects, breaks = seq(0, max(num_projects), 2))) %>%
  summarise(num_projects = sum(num_projects)) %>% 
  ungroup()
```

```{r}
dt %>%
  filter(total_donations < 2000) %>%
  ggplot(aes(x=total_donations)) +
  geom_histogram() +
  theme_classic()
```

# binwidth
```{r}
dt %>%
  filter(total_donations < 2000) %>%
  ggplot(aes(x=total_donations)) +
  geom_histogram(binwidth = 200) +
  theme_classic()
```

# bins
```{r}
dt %>%
  filter(total_donations < 2000) %>%
  ggplot(aes(x=total_donations)) +
  geom_histogram(bins = 5) +
  theme_classic()
```

# Cumulative Frequency Distribution
```{r}
dt %>%
  filter(total_donations < 2000) %>%
  mutate(bins = cut(total_donations, breaks = 10, dig.lab = 4)) %>%
  group_by(bins) %>%
  summarise(count = n()) %>% 
  arrange(bins) %>%
  mutate(cum_freq = cumsum(count)) %>%
  ungroup()

#generate a column called bin (cut() is used to convert Numeric to Factor)
#breaks = 10 give the number of intervals
#dig.lab: It determines the number of digits used in formatting the break numbers.
#cumsum() Returns a number which is the cumulative sums.
```
# Cumulative Relative Frequency Distribution
```{r}
dt %>%
  filter(total_donations < 2000) %>%
  mutate(bins = cut(total_donations, breaks = 10, dig.lab = 4)) %>%
  group_by(bins) %>%
  summarise(count = n()) %>%
  arrange(bins) %>%
  mutate(cum_freq = cumsum(count), rel_cum_freq = cum_freq/sum(count)) %>%
  ungroup()
```

```{r}
dt %>%
  filter(total_donations < 2000) %>%
  mutate(bins = cut(total_donations, breaks = 10, dig.lab = 4)) %>%
  group_by(bins) %>%
  summarise(count = n()) %>%
  arrange(bins) %>%
  mutate(cum_freq = cumsum(count), rel_cum_freq = cum_freq/sum(count)) %>%
  ungroup() %>%
  ggplot(aes(x=bins,y=rel_cum_freq, group=1)) +
  geom_point() +
  geom_line() +
  theme_classic()
```


```{r}
dt1 <- read_csv("projects.csv")

```


## Pie Charts

```{r}

dt1 %>% 
  select(id, poverty_level, total_price_excluding_optional_support) %>%
  group_by(poverty_level) %>%
  summarise(n = n()) %>% 
  ungroup() %>% 
  ggplot(aes(x="Pie Chart of Poverty Level", y=n, fill=poverty_level)) +
  geom_bar(width=1, stat="identity")+
  coord_polar("y", start=0) +
  theme_classic()
```

# bar charts using ggplot
```{r}
dt1 %>% 
  select(id, poverty_level) %>%
  ggplot(aes(x=poverty_level, fill=poverty_level)) +
  geom_bar() +
  theme_classic ()
```



## Grouped Bar Charts
```{r}
dt1 %>% 
  select(id, poverty_level, funding_status) %>%
  ggplot(aes(x=poverty_level, fill=poverty_level)) +
  geom_bar() +
  facet_grid(. ~ funding_status) +
  theme_classic()
```

## Stacked Bar Chart
```{r}
dt1 %>% 
  select(id, poverty_level, funding_status) %>%
  group_by(poverty_level, funding_status) %>%
  summarise(n = n()) %>%
  ungroup() %>% 
  ggplot(aes(x=funding_status, y=n, fill=poverty_level)) +
  geom_bar(stat="identity") +
  theme_classic()
```

## Horizontal Grouped Barchart
```{r}
dt1 %>% 
    select(id, poverty_level, funding_status) %>%
  ggplot(aes(x=poverty_level, fill=poverty_level)) +
  geom_bar() +
  facet_grid(. ~ funding_status) +
  coord_flip() +
  theme_classic()
```


##Constructing line chart
```{r}
myData3 <- read_xlsx("Class.xlsx","Apple_Merck")
plot(myData3$Apple ~ myData3$Date, main="Monthly stock prices for Apple and Merck", xlab="Date", ylab="Monthly stock price (in $)",
     col = "blue", type = "l", ylim=c(0,300))
lines(myData3$Merck~myData3$Date, col="red", type = "l")
legend("topleft", legend=c("Apple","Merck"), col=c("blue","red"), lty=1)

```
## Line Charts using ggplot
```{r}
dt %>%
  select(total_price_excluding_optional_support, date_posted) %>%
  mutate(year = year(mdy(date_posted))) %>%
  group_by(year) %>%
  summarize(avg_size = mean(total_price_excluding_optional_support)) %>%
  ungroup() %>% 
  ggplot(aes(year, avg_size)) +
  geom_line() +
  theme_classic()
```

## Grouped Line Charts (1)
```{r}
dt %>%
  select(total_price_excluding_optional_support, poverty_level, date_posted) %>%
  mutate(year = year(mdy(date_posted))) %>%
  group_by(year, poverty_level) %>%
  summarize(avg_size = mean(total_price_excluding_optional_support)) %>%
  ungroup() %>% 
  ggplot(aes(year, avg_size, colour=poverty_level)) +
  geom_line() +
  theme_classic()
```

## Grouped Line Charts (2)
```{r}
dt %>%
  select(total_price_excluding_optional_support, poverty_level, date_posted) %>%
  mutate(year = year(mdy(date_posted))) %>%
  group_by(year, poverty_level) %>%
  summarize(avg_size = mean(total_price_excluding_optional_support)) %>%
  ungroup() %>% 
  filter(poverty_level=="high"|poverty_level=="unknown") %>% ggplot(aes(year, avg_size, colour=poverty_level)) +
  geom_line() +
  theme_classic()
```

## Grouped Line Charts (3)
```{r}
dt %>%
  select(total_price_excluding_optional_support, poverty_level, date_posted) %>%
  mutate(year = year(mdy(date_posted))) %>%
  group_by(year, poverty_level) %>%
  summarize(avg_size = mean(total_price_excluding_optional_support)) %>%
  ungroup() %>% 
  filter(poverty_level=="low"|poverty_level=="minimal") %>% ggplot(aes(year, avg_size, colour=poverty_level)) +
  geom_line() +
  theme_classic()
```

## Area Charts
```{r}
dt %>%
  select(total_price_excluding_optional_support, total_price_including_optional_support, date_posted) %>%
  mutate(year = year(mdy(date_posted))) %>%
  group_by(year) %>%
  summarize(total_size = sum(total_price_excluding_optional_support),
            total_optional = sum(total_price_including_optional_support)) %>%
  gather("var", "value", -year) %>%
  ggplot(aes(year, value, fill=var)) +
  geom_area(alpha=0.8) +
  theme_classic()
```

## Scatter Charts using ggplot
```{r}
dt %>%
  sample_n(50) %>%
  ggplot(aes(total_price_excluding_optional_support, num_donors)) +
  geom_point() +
  theme_classic()
```


#Constructing bubble plot
```{r}
myData2 <- read_xlsx("Class.xlsx","Birth")
plot(myData2$Birth_Rate ~ myData2$Life_Exp, type="n")
symbols(myData2$Birth_Rate ~ myData2$Life_Exp, circles=myData2$GNI, inches=0.5, bg="blue", main="A bubble plot of Birth Rate, Life Expectancy, and GNI", xlab = "Life Expectancy (in years)", ylab ="Birth Rate (in %)")

```
## Bubble Charts using ggplot
```{r}
dt %>%
  sample_n(50) %>%
  ggplot(aes(total_price_excluding_optional_support, num_donors, size=num_donors)) +
  geom_point() +
  theme_classic()
```

#Constructing heatmap
```{r}
#Import the Bookstores data file and label it myData
myData3 <- read_xlsx("Class.xlsx","Bookstore")
myTable <-table(myData3$BookStore, myData3$BookType)
myTable <- myTable/rowSums(myTable)
myData3.matrix <- as.matrix(myTable)
heatmap(myData3.matrix, col = heat.colors(256), scale ="none", cexCol=0.75, cexRow=1,
        Rowv= NA, Colv = NA)

myBreaks <- c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30)
myCol <- c("blue", "green", "yellow", "orange1", "red1",
           "red3")
heatmap(myData3.matrix, col = myCol, breaks = myBreaks, scale ="none", cexCol=0.75, cexRow=1, Rowv = NA, Colv = NA)

```




```{r}
# Scatter plot with additional options
library(ggplot2)
ggplot(data=mtcars, aes(x=wt, y=mpg)) +
  geom_point(pch=17, color="blue", size=2) +
  geom_smooth(method="lm", color="red", linetype=2) +
  labs(title="Automobile Data", x="Weight", y="Miles Per Gallon")
```


```{r}
# Scatter plot with faceting and grouping
data(mtcars)
mtcars$am <- factor(mtcars$am, levels=c(0,1),
                    labels=c("Automatic", "Manual"))
mtcars$vs <- factor(mtcars$vs, levels=c(0,1),
                    labels=c("V-Engine", "Straight Engine"))
mtcars$cyl <- factor(mtcars$cyl)
```


```{r}
library(ggplot2)
ggplot(data=mtcars, aes(x=hp, y=mpg,
                        shape=cyl, color=cyl)) +
  geom_point(size=3) +
  facet_grid(am~vs) +
  labs(title="Automobile Data by Engine Type",
       x="Horsepower", y="Miles Per Gallon")
```


```{r}
# Using geoms
data(singer, package="lattice")
ggplot(singer, aes(x=height)) + geom_histogram()

ggplot(singer, aes(x=voice.part, y=height)) + geom_boxplot()

data(Salaries, package="car")
library(ggplot2)
ggplot(Salaries, aes(x=rank, y=salary)) +
  geom_boxplot(fill="cornflowerblue",
               color="black", notch=TRUE)+
  geom_point(position="jitter", color="blue", alpha=.5)+
  geom_rug(side="l", color="black")
```


```{r}
# Grouping
library(ggplot2)
data(singer, package="lattice")
ggplot(singer, aes(x=voice.part, y=height)) +
  geom_violin(fill="lightblue") +
  geom_boxplot(fill="lightgreen", width=.2)
```


```{r}
data(Salaries, package="car")
library(ggplot2)
ggplot(data=Salaries, aes(x=salary, fill=rank)) +
  geom_density(alpha=.3)

ggplot(Salaries, aes(x=yrs.since.phd, y=salary, color=rank,
                     shape=sex)) + geom_point()

ggplot(Salaries, aes(x=rank, fill=sex)) +
  geom_bar(position="stack") + labs(title='position="stack"')

ggplot(Salaries, aes(x=rank, fill=sex)) +
  geom_bar(position="dodge") + labs(title='position="dodge"')

ggplot(Salaries, aes(x=rank, fill=sex)) +
  geom_bar(position="fill") + labs(title='position="fill"')
```


```{r}
# Placing options
ggplot(Salaries, aes(x=rank, fill=sex))+ geom_bar()

ggplot(Salaries, aes(x=rank)) + geom_bar(fill="red")

ggplot(Salaries, aes(x=rank, fill="red")) + geom_bar()
```


```{r}
# Faceting
data(singer, package="lattice")
library(ggplot2)
ggplot(data=singer, aes(x=height)) +
  geom_histogram() +
  facet_wrap(~voice.part, nrow=4)
```


```{r}
library(ggplot2)
ggplot(Salaries, aes(x=yrs.since.phd, y=salary, color=rank,
                     shape=rank)) + geom_point() + facet_grid(.~sex)

data(singer, package="lattice")
library(ggplot2)
ggplot(data=singer, aes(x=height, fill=voice.part)) +
  geom_density() +
  facet_grid(voice.part~.)
```


```{r}
# Adding smoothed lines
data(Salaries, package="car")
library(ggplot2)
ggplot(data=Salaries, aes(x=yrs.since.phd, y=salary)) +
  geom_smooth() + geom_point()

ggplot(data=Salaries, aes(x=yrs.since.phd, y=salary,
                          linetype=sex, shape=sex, color=sex)) +
  geom_smooth(method=lm, formula=y~poly(x,2),
              se=FALSE, size=1) +
  geom_point(size=2)
```


```{r}
# Modifying axes
data(Salaries,package="car")
library(ggplot2)
ggplot(data=Salaries, aes(x=rank, y=salary, fill=sex)) +
  geom_boxplot() +
  scale_x_discrete(breaks=c("AsstProf", "AssocProf", "Prof"),
                   labels=c("Assistant\nProfessor",
                            "Associate\nProfessor",
                            "Full\nProfessor")) +
  scale_y_continuous(breaks=c(50000, 100000, 150000, 200000),
                     labels=c("$50K", "$100K", "$150K", "$200K")) +
  labs(title="Faculty Salary by Rank and Sex", x="", y="")
```


```{r}
# Legends
data(Salaries,package="car")
library(ggplot2)
ggplot(data=Salaries, aes(x=rank, y=salary, fill=sex)) +
  geom_boxplot() +
  scale_x_discrete(breaks=c("AsstProf", "AssocProf", "Prof"),
                   labels=c("Assistant\nProfessor",
                            "Associate\nProfessor",
                            "Full\nProfessor")) +
  scale_y_continuous(breaks=c(50000, 100000, 150000, 200000),
                     labels=c("$50K", "$100K", "$150K", "$200K")) +
  labs(title="Faculty Salary by Rank and Gender",
       x="", y="", fill="Gender") +
  theme(legend.position=c(.1,.8))
```


```{r}
# Scales
ggplot(mtcars, aes(x=wt, y=mpg, size=disp)) +
  geom_point(shape=21, color="black", fill="cornsilk") +
  labs(x="Weight", y="Miles Per Gallon",
       title="Bubble Chart", size="Engine\nDisplacement")
```


```{r}
data(Salaries, package="car")
ggplot(data=Salaries, aes(x=yrs.since.phd, y=salary, color=rank)) +
  scale_color_manual(values=c("orange", "olivedrab", "navy")) +
  geom_point(size=2)

ggplot(data=Salaries, aes(x=yrs.since.phd, y=salary, color=rank)) +
  scale_color_brewer(palette="Set1") + geom_point(size=2)

library(RColorBrewer)
display.brewer.all()
```


```{r}
# Themes
data(Salaries, package="car")
library(ggplot2)
mytheme <- theme(plot.title=element_text(face="bold.italic",
                                         size="14", color="brown"),
                 axis.title=element_text(face="bold.italic",
                                         size=10, color="brown"),
                 axis.text=element_text(face="bold", size=9,
                                        color="darkblue"),
                 panel.background=element_rect(fill="white",
                                               color="darkblue"),
                 panel.grid.major.y=element_line(color="grey",
                                                 linetype=1),
                 panel.grid.minor.y=element_line(color="grey",
                                                 linetype=2),
                 panel.grid.minor.x=element_blank(),
                 legend.position="top")

ggplot(Salaries, aes(x=rank, y=salary, fill=sex)) +
  geom_boxplot() +
  labs(title="Salary by Rank and Sex", 
       x="Rank", y="Salary") +
  mytheme
```


```{r}
# Multiple graphs per page
data(Salaries, package="car")
library(ggplot2)
p1 <- ggplot(data=Salaries, aes(x=rank)) + geom_bar()
p2 <- ggplot(data=Salaries, aes(x=sex)) + geom_bar()
p3 <- ggplot(data=Salaries, aes(x=yrs.since.phd, y=salary)) + geom_point()

library(gridExtra)
grid.arrange(p1, p2, p3, ncol=3)
```


```{r}
# Saving graphs
ggplot(data=mtcars, aes(x=mpg)) + geom_histogram()
ggsave(file="mygraph.pdf")

```

