---
title: "Week 11 Materials"
output:
  html_document:
    df_print: paged
---
```{r}
#install.packages("caret", dependencies = c("Depends", "Suggests"))
#install.packages("gains")
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("pROC")
library(caret)
library(gains)
library(rpart)
library(rpart.plot)
library(pROC)
library(readxl)
```
#Load data and transform data
#set the random seed to 1, thus generating the same partitions as in this example
```{r}
myData<-read_excel("Week_11.xlsx", sheet = "HELOC_Data")
myData$HELOC <- as.factor(myData$HELOC)
set.seed(1)
myIndex <- createDataPartition(myData$HELOC, p=0.7, list=FALSE)
trainSet <- myData[myIndex,]
validationSet <- myData[-myIndex,]

```

#We use the rpart function to generate the default classification tree, default_tree. Within the rpart function, we specify the model structure, data source, and method. The method option is set to “class” for developing a classification tree. To view the details about the default tree, use the summary function. Because R uses the k-fold cross-validation method for pruning the tree, to ensure consistency of the cross-validation results, we use the set.seed function to set a random seed of 1
#use the prp function to view the classification tree visually
```{r}
set.seed(1)
default_tree <- rpart(HELOC ~., data = trainSet, method = "class")
summary(default_tree)
prp(default_tree, type = 1, extra = 1, under = TRUE)
```

#grow the full tree by using the rpart function.

```{r}
set.seed(1)
full_tree <- rpart(HELOC ~ ., data = trainSet, method = "class", cp = 0, minsplit = 2, minbucket = 1)
prp(full_tree, type = 1, extra = 1, under = TRUE)
printcp(full_tree) #display the complexity parameter table
```

#prune
```{r}
pruned_tree <- prune(full_tree, cp = 0.0166)
prp(pruned_tree, type = 1, extra = 1, under = TRUE)
predicted_class <- predict(pruned_tree, validationSet, type = "class")
confusionMatrix(predicted_class, validationSet$HELOC, positive = "1")
```

#predict probability and change the cut-off
```{r}
predicted_prob <- predict(pruned_tree, validationSet, type= 'prob')
head(predicted_prob)
confusionMatrix(as.factor(ifelse(predicted_prob[,2]>0.26, '1', '0')), validationSet$HELOC, positive = '1')
validationSet$HELOC <- as.numeric(as.character(validationSet$HELOC))
```

#Generating cumulative lift plots
```{r}
gains_table <- gains(validationSet$HELOC, predicted_prob[,2])
gains_table
plot(c(0, gains_table$cume.pct.of.total*sum(validationSet$HELOC)) ~ c(0, gains_table$cume.obs), xlab = '# of cases', ylab = "Cumulative", type = "l")
lines(c(0, sum(validationSet$HELOC))~c(0, dim(validationSet)[1]), col="red", lty=2)
barplot(gains_table$mean.resp/mean(validationSet$HELOC), names.arg=gains_table$depth, xlab="Percentile", ylab="Lift", ylim=c(0, 3.0), main="Decile-Wise Lift Chart")
roc_object <- roc(validationSet$HELOC, predicted_prob[,2])
plot.roc(roc_object)
auc(roc_object)
```

#Predict new cases
```{r}
myScoreData<-read_excel("Week_11.xlsx", sheet = "HELOC_Score")
predicted_class_score <- predict(pruned_tree, myScoreData, type = "class")
predicted_class_score
predicted_class_prob <- predict(pruned_tree, myScoreData, type = "prob")
predicted_class_prob
```

#Example 13.7
#Import the data from the Balance_Data worksheet of the Balance data file into a data frame (table) and label it myData.
#If you are using a newer version of R than version 3.5.3, execute the following line of code:
```{r}
#install.packages("caret", dependencies = c("Depends", "Suggests"))
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("pROC")
#install.packages("forecast")
```



```{r}
myData<-read_excel("Week_11.xlsx", sheet = "Balance_Data")
library(rpart.plot)
library(forecast)
```
#Cross-validation
```{r}
set.seed(1)
myIndex <- createDataPartition(myData$Balance, p=0.7, list=FALSE)
trainSet <- myData[myIndex,]
validationSet <- myData[-myIndex,]
set.seed(1)
```
#default tree
```{r}
default_tree <- rpart(Balance ~., data = trainSet, method = "anova")
summary(default_tree)
prp(default_tree, type = 1, extra = 1, under = TRUE)
```

#full tree and prune
```{r}
set.seed(1)
full_tree <- rpart(Balance ~ ., data = trainSet, method = "anova", cp = 0, minsplit = 2, minbucket = 1)
prp(full_tree, type = 1, extra = 1, under = TRUE)
printcp(full_tree)
pruned_tree <- prune(full_tree, cp = 7.5219e-02)
prp(pruned_tree, type = 1, extra = 1, under = TRUE)
predicted_value <- predict(pruned_tree, validationSet)
accuracy(predicted_value, validationSet$Balance)
```

#predict new
```{r}
myScoreData<-read_excel("Week_11.xlsx", sheet = "Balance_Score")
predicted_value_score <- predict(pruned_tree, myScoreData)
predicted_value_score
```

#Example 13.8 randomForest
```{r}
#install.packages("randomForest")
library(caret)
library(gains)
library(pROC)
library(randomForest)
```
#cross-validation
```{r}
myData<-read_excel("Week_11.xlsx", sheet = "HELOC_Data")
myData$HELOC <- as.factor(myData$HELOC)
myData$Sex<-as.factor(myData$Sex)
set.seed(1)
myIndex <- createDataPartition(myData$HELOC, p=0.6, list=FALSE)
trainSet <- myData[myIndex,]
validationSet <- myData[-myIndex,]
set.seed(1)
```

#RandomForest：Bagging tree 装袋减少方差
#step e
```{r}
bagging_tree <- randomForest(HELOC ~., data = trainSet, ntree = 100, mtry = 3, importance = TRUE)
varImpPlot(bagging_tree, type=1)
predicted_class <- predict(bagging_tree, validationSet)
confusionMatrix(predicted_class, validationSet$HELOC, positive = "1")
```

#Cumulative Lift
```{r}
predicted_prob <- predict(bagging_tree, validationSet, type= 'prob')
validationSet$HELOC <- as.numeric(as.character(validationSet$HELOC))
gains_table <- gains(validationSet$HELOC, predicted_prob[,2])
gains_table
plot(c(0, gains_table$cume.pct.of.total*sum(validationSet$HELOC)) ~ c(0, gains_table$cume.obs), xlab = '# of cases', ylab = "Cumulative", type = "l")
lines(c(0, sum(validationSet$HELOC))~c(0, dim(validationSet)[1]), col="red", lty=2)
barplot(gains_table$mean.resp/mean(validationSet$HELOC), names.arg=gains_table$depth, xlab="Percentile", ylab="Lift", ylim=c(0, 3.0), main="Decile-Wise Lift Chart")
roc_object <- roc(validationSet$HELOC, predicted_prob[,2])
plot.roc(roc_object)
auc(roc_object)
```
#Predict New 
```{r}
myScoreData<-read_excel("Week_11.xlsx", sheet = "HELOC_Score")
myScoreData$Sex <- as.factor(myScoreData$Sex)
predicted_class_score <- predict(bagging_tree, myScoreData, type = "class")
predicted_class_score
predicted_class_prob <- predict(bagging_tree, myScoreData, type = "prob")
predicted_class_prob
```



#random forest. Replace the R commands in step e above using the following commands:
```{r}
set.seed(1)
randomforest_tree <- randomForest(HELOC ~., data = trainSet, ntree = 100, mtry = 2, importance = TRUE)
varImpPlot(randomforest_tree, type=1)
predicted_class <- predict(randomforest_tree, validationSet)
confusionMatrix(predicted_class, as.factor(validationSet$HELOC), positive = "1")
predicted_prob <- predict(randomforest_tree, validationSet, type= 'prob')
```


#boosting
```{r}
#install.packages("adabag")
library(caret)
library(pROC)
library(gains)
library(adabag)

myData <- data.frame(myData)
myData$HELOC <- as.factor(myData$HELOC)
myData$Sex<-as.factor(myData$Sex)
set.seed(1)
myIndex <- createDataPartition(myData$HELOC, p=0.6, list=FALSE)
trainSet <- myData[myIndex,]
validationSet <- myData[-myIndex,]
set.seed(1)
```


```{r}
boosting_tree <- boosting(HELOC ~ ., data = trainSet, mfinal = 100)
prediction <- predict(boosting_tree, validationSet)
confusionMatrix(as.factor(prediction$class), validationSet$HELOC, positive = "1")
validationSet$HELOC <- as.numeric(as.character(validationSet$HELOC))
```


```{r}
gains_table <- gains(validationSet$HELOC, prediction$prob[,2])
gains_table
plot(c(0, gains_table$cume.pct.of.total*sum(validationSet$HELOC)) ~ c(0, gains_table$cume.obs), xlab = '# of cases', ylab = "Cumulative", type = "l")
lines(c(0, sum(validationSet$HELOC))~c(0, dim(validationSet)[1]), col="red", lty=2)
barplot(gains_table$mean.resp/mean(validationSet$HELOC), names.arg=gains_table$depth, xlab="Percentile", ylab="Lift", ylim=c(0, 3.0), main="Decile-Wise Lift Chart")
roc_object <- roc(validationSet$HELOC, prediction$prob[,2])
plot.roc(roc_object)
auc(roc_object)
```


```{r}
myScoreData <- data.frame(myScoreData)
myScoreData$Sex <- as.factor(myScoreData$Sex)
predicted_class_score <- predict(boosting_tree, myScoreData)
predicted_class_score$class
predicted_class_score$prob
```
