---
title: "Week 5 Material"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
library(tidyverse)
library(readxl)
library(knitr)
library(psych)
options("scipen"=100, "digits"=4)
```

## if the above is not working:
```{r}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
```

# Computing Probabilities with Standard Normal Tables
# The distribution for customer demand (units per month) is normal with: mean=750 stdev.=100
# (a) what is the probability  that demand will be at least 900 units/month?
```{r}
pnorm(900, 750, 100)
```
# (b) Probability that demand will exceed 700 units, or P(X>700).
```{r}
1-pnorm(700,750,100)
```

# (c) Probability that demand will be between 700 and 900, or P(700<X<900).
```{r}
pnorm(900,750,100)-pnorm(700,750,100)
```

# Examples of Using R to plot probability distribution
# 1. Plot the standard normal curve on the interval [–3,3] 
```{r}
# seq Generate regular sequences.
x <- seq(from = -3, to = 3, by = 0.1)

y = dnorm(x)
data <- data.frame(x = x, y=y)
ggplot(data, aes(x, y)) +
geom_line() +
labs(x = "Normal Deviate",
y = "Density") +
# A function that takes the limits as input and returns breaks as output 
scale_x_continuous(breaks = seq(-3, 3, 1))
```
# 2. What is the area under the standard normal curve to the left of z=1.96?
```{r}
#by default mean=0 sd=1
pnorm(1.96,mean=0,sd=1)
```
# 3. What is the value of the 90th percentile of a normal distribution with a mean of 500 and a standard deviation of 100?
```{r}
qnorm(.9, mean=500, sd=100)
```

# 4. Generate 50 random normal deviates with a mean of 50 and a standard deviation of 10.
# Each time you generate pseudo-random deviates, a different seed, and therefore different results, are produced. To make your results reproducible, you can specify the seed explicitly, using the set.seed()
```{r}
set.seed(1234) 
x <- rnorm(50, mean=50, sd=10)
x
```

# Analyzing Airline Passenger Data
```{r}
air<-read_excel("Airline Passengers.xlsx", range="Data!A3:A28")
```
```{r}
hist(air$`Sample Data`)
```
```{r}
describe(air$`Sample Data`)
```
# Goodness of Fit
```{r}
dt <- read_csv("projects.csv")
```

```{r}
dt %>%
filter(total_price_excluding_optional_support<2000) %>%
ggplot(aes(x=total_price_excluding_optional_support)) +
  geom_histogram() +
  theme_classic()
```
```{r}
dt %>%
  filter(total_price_excluding_optional_support < 2000) %>%
  ggplot(aes(total_price_excluding_optional_support)) +
  geom_density() +
  theme_classic()
```

检测样本是否来源于服从某个特定理论概率分布的总体(population）。原理是将empirical cumulative distribution function  (ECDF) 和任意理论分布的(CDF)进行比较。只适用于连续变量。

原假设H0: 样本来自于给定的理论概率分布 The data follow the specific distribution

备择假设H1：样本不属于理论的概率分布 The data do not follow the specific distribution

```{r}
ks.test(dt$total_price_excluding_optional_support,"pnorm")
```
```{r}
shapiro.test(sample(dt$total_price_excluding_optional_support, 5000))
```
#we reject the null hypothesis. We have sufficient evidence to say that the sample data does not come from a normal distribution.



```{r}
dt %>%
  filter(total_price_excluding_optional_support < 2000) %>%
  ggplot(aes(total_price_excluding_optional_support)) +
  geom_density() +
  theme_classic()
```


```{r}
dt %>%
  filter(total_price_excluding_optional_support < 2000) %>%
  ggplot(aes(sample=total_price_excluding_optional_support)) +
  stat_qq() +
  stat_qq_line() +
  theme_classic()
```


# Sampling experiment in R
```{r}
set.seed(12345)

sample10 <- matrix(ncol=25, nrow=10)
sample25 <- matrix(ncol=25, nrow=25)
sample100 <- matrix(ncol=25, nrow=100)
sample500 <- matrix(ncol=25, nrow=500)

for(i in 1:25){
  sample10[,i] <- runif(10,0,10)
  sample25[,i] <- runif(25,0,10)
  sample100[,i] <- runif(100,0,10)
  sample500[,i] <- runif(500,0,10)
}
```


```{r}
hist(sample10)
hist(colMeans(sample10))
mean(colMeans(sample10))
sd(colMeans(sample10))
mean(colMeans(sample10))-sd(colMeans(sample10))*3
mean(colMeans(sample10))+sd(colMeans(sample10))*3
(mean(colMeans(sample10))+sd(colMeans(sample10))*3)-(mean(colMeans(sample10))-sd(colMeans(sample10))*3)
```

```{r}
hist(sample25)
hist(colMeans(sample25))
mean(colMeans(sample25))
sd(colMeans(sample25))
mean(colMeans(sample25))-sd(colMeans(sample25))*3
mean(colMeans(sample25))+sd(colMeans(sample25))*3
(mean(colMeans(sample25))+sd(colMeans(sample25))*3)-(mean(colMeans(sample25))-sd(colMeans(sample25))*3)
```


```{r}
hist(sample100)
hist(colMeans(sample100))
mean(colMeans(sample100))
sd(colMeans(sample100))
mean(colMeans(sample100))-sd(colMeans(sample100))*3
mean(colMeans(sample100))+sd(colMeans(sample100))*3
(mean(colMeans(sample100))+sd(colMeans(sample100))*3)-(mean(colMeans(sample100))-sd(colMeans(sample100))*3)
```

```{r}
hist(sample500)
hist(colMeans(sample500))
mean(colMeans(sample500))
sd(colMeans(sample500))
mean(colMeans(sample500))-sd(colMeans(sample500))*3
mean(colMeans(sample500))+sd(colMeans(sample500))*3
(mean(colMeans(sample500))+sd(colMeans(sample500))*3)-(mean(colMeans(sample500))-sd(colMeans(sample500))*3)
```

# Confidence Interval for the Mean with Known Population Standard Deviation
```{r}
796 + qnorm(1-(0.05)/2) * (15/sqrt(25))
796 - qnorm(1-(0.05)/2) * (15/sqrt(25))
```

# Confidence Interval for the Mean with  Unknown Population Standard Deviation(1)
```{r}
qt(1-(0.05)/2, df=29)
qt(1-(0.05), df=29)

qt(1-(0.05)/2, df=99)
qt(1-(0.05), df=99)

qt(1-(0.05)/2, df=999)
qt(1-(0.05), df=999)

qnorm(1-(0.05)/2)
qnorm(1-(0.05))
```


# Confidence Interval for the Mean with  Unknown Population Standard Deviation(2)
```{r}
dt2 <- read_excel("Credit Approval Decisions.xlsx",skip=2, sheet="Data")
mean(dt2$`Revolving Balance`[dt2$Homeowner=="Y"])
sd(dt2$`Revolving Balance`[dt2$Homeowner=="Y"])
nrow(dt2[dt2$Homeowner=="Y",])

mean(dt2$`Revolving Balance`[dt2$Homeowner=="Y"]) + qt(1-(0.05)/2, 26) * (sd(dt2$`Revolving Balance`[dt2$Homeowner=="Y"])/sqrt(27))
mean(dt2$`Revolving Balance`[dt2$Homeowner=="Y"]) - qt(1-(0.05)/2, 26) * (sd(dt2$`Revolving Balance`[dt2$Homeowner=="Y"])/sqrt(27))
```
```