---
title: "Week 6 Materials"
output:
  html_document:
    df_print: paged
---
```{r}
library(tidyverse)
library(readxl)
```

## if the above is not working:
```{r}
library(dplyr)
library(readxl)
```

# 7.4
```{r}
ca <- read_excel("CadSoft Technical Support Response Times.xlsx", skip=1)
```
```{r}
t.test(ca$`Time (min)`,mu=25,alternative="less")
```
# 7.6
```{r}
vc <- read_excel("Vacation Survey.xlsx", skip=1)
```
```{r}
t.test(vc$Age, mu=35, alternative = "two.sided")
```
# 7.9
```{r}
ps <- read_excel("Purchase Orders.xlsx", skip=1)
```

```{r}
ps <- ps %>%
  mutate(lead_time=`Arrival Date` - `Order Date`)
```
  
```{r}
ps %>%
	group_by(Supplier) %>%
	summarise(average_lead_time = mean(lead_time))

```

```{r}
t.test(lead_time ~ Supplier, 
       data = ps %>% 
         filter(Supplier == "Alum Sheeting" | Supplier == "Durrable Products"),
       alternative="greater")
```

```{r}
t.test(lead_time ~ Supplier, 
       data = ps %>% 
         filter(Supplier == "Alum Sheeting" | Supplier == "Durrable Products"),
       alternative="less")
```

# 7.11 paired t-test
```{r}
pf <- read_excel("Pile Foundation.xlsx", skip=3)
```

```{r}
t.test(pf$Estimated, pf$Actual, paired=TRUE)
```

# 7.12 F-Test for Equality of Variances  
```{r}
var.test(lead_time ~ Supplier, 
       data = ps %>% 
         filter(Supplier == "Alum Sheeting" | Supplier == "Durrable Products"),
       alternative="less")
```

# 7.13 read and check the data
```{r}
is <- read_excel("Insurance Survey.xlsx", skip=2, n_max=25)
head (is)
str(is)
# check the classifications of variables 
```

# One-way ANOVA
# aov(formula, data = NULL) we use this function to run the model

# One-way ANOVA (manual)
```{r}
mean(is$Satisfaction[is$Education=="College graduate"])
mean(is$Satisfaction[is$Education=="Graduate degree"])
mean(is$Satisfaction[is$Education=="Some college"])

var(is$Satisfaction[is$Education=="College graduate"])
var(is$Satisfaction[is$Education=="Graduate degree"])
var(is$Satisfaction[is$Education=="Some college"])

# Grand mean = 3.708
mean(mean(is$Satisfaction[is$Education=="College graduate"])*9/24+mean(is$Satisfaction[is$Education=="Graduate degree"])*8/24+mean(is$Satisfaction[is$Education=="Some college"])*7/24)

# SSA
9*(3.444-3.708)^2+8*(4.500-3.708)^2+7*(3.143-3.708)^2

# MSA
(9*(3.444-3.708)^2+8*(4.500-3.708)^2+7*(3.143-3.708)^2)/(3-1)

# SSW
(9-1)*1.028+(8-1)*0.571+(7-1)*1.476

# MSW
((9-1)*1.028+(8-1)*0.571+(7-1)*1.476)/(24-3)

# SST
(9*(3.444-3.708)^2+8*(4.500-3.708)^2+7*(3.143-3.708)^2)+((9-1)*1.028+(8-1)*0.571+(7-1)*1.476)

# F-test statistic
((9*(3.444-3.708)^2+8*(4.500-3.708)^2+7*(3.143-3.708)^2)/(3-1)) / (((9-1)*1.028+(8-1)*0.571+(7-1)*1.476)/(24-3))

# F-test
qf(0.05, 2, 21, lower.tail = FALSE)
pf(3.925, 2, 21, lower.tail = FALSE)

# R2
(9*(3.444-3.708)^2+8*(4.500-3.708)^2+7*(3.143-3.708)^2)/((9*(3.444-3.708)^2+8*(4.500-3.708)^2+7*(3.143-3.708)^2)+((9-1)*1.028+(8-1)*0.571+(7-1)*1.476))

1-(((9-1)*1.028+(8-1)*0.571+(7-1)*1.476)/((9*(3.444-3.708)^2+8*(4.500-3.708)^2+7*(3.143-3.708)^2)+((9-1)*1.028+(8-1)*0.571+(7-1)*1.476)))
```
# One-way ANOVA in R
```{r}
aov<-aov(Satisfaction~Education, data = is)
```

```{r}
summary(aov)
#summary(aov(is$Satisfaction~is$Education))
```
# Tukey's HSD
```{r}
TukeyHSD(aov)
```

# One-way ANOVA in a linear model (reference-coded model)
```{r}
lm<-lm(Satisfaction~Education, data = is)
#lm<-lm(Satisfaction~as.factor(Education), data = is)
```
```{r}
summary(lm)
```

# Check for homoscedasticity (homogeneity of variance)
# For the that follows, the null hypothesis is that all populations variances are equal, the alternative hypothesis is that at least two of them differ. Consequently, p-values less than 0.1, 0.05, 0.001 (depending on your desired threshold) suggest variances are significantly different and the homogeneity of variance assumption has been violated.
```{r}
bartlett.test(Satisfaction~Education, data = is)
```

