---
title: "Week 3 Tutorial and Bonus"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r}
library(tidyverse)
library(knitr)
options("scipen"=100, "digits"=4)
```

```{r}
rm(list = ls())
dt <- read_csv("projects.csv")
```

```{r}
dt %>%
  select(id, grade_level, poverty_level)
```
# Contingency Table 列联表

```{r}
dt %>%
  select(grade_level, poverty_level) %>%
  table()
```

```{r}
dt %>%
  select(grade_level, poverty_level) %>%
  table() %>%
  prop.table()
```

```{r}
dt %>%
  select(poverty_level, resource_type, grade_level) %>%
  table()
```

```{r}
dt %>%
  select(grade_level, poverty_level) %>%
  filter(poverty_level != "unknown") 
```

# Pivot Table
```{r}
library(rpivotTable)
dt %>% 
  filter(poverty_level != "unknown")  %>%
  filter(is.na(grade_level)==FALSE)    %>%
  rpivotTable(rows=c("grade_level"), cols=c("poverty_level"))
```

# Measures of location: Application
```{r}
library(modeest)
```


# Measures of location: Application - Repair time
```{r}
library(readxl)
repair<-read_excel("Computer Repair Times.xlsx", skip=2,sheet=1)
repair <- repair %>% rename(time = `Repair Time (Days)`)
```


```{r}
repair %>% 
  ggplot(aes(x=time)) +
  geom_histogram(binwidth = 1) +
  theme_classic()
```
```{r}
repair %>%
  group_by(time) %>%
  summarise(count = n()) %>% 
  arrange(time) %>%
  mutate(cum_freq = cumsum(count)) %>%
  mutate(cum_freq = cumsum(count), rel_cum_freq = cum_freq/sum(count)) %>%
  ungroup()
```

# Measures of disperson - variance and standard deviation
```{r}
#library(readxl)
stock<-read_excel("Closing Stock Prices.xlsx",skip=2)
head(stock) # Name of the last column contains special characters, but this only can be checked in R console!
stock<- stock %>% rename(`DJ Industrials Index`=`DJ Industrials \r\nIndex`)
```


# Computing z-Score
```{r}
dt2 <- stock %>% mutate(`IBM: z-score`=(`IBM`-mean(`IBM`))/sd(`IBM`))
head(dt2)
```

# Skewness
```{r}
mean((dt2$`IBM`-mean(dt2$`IBM`))^3)/sd(dt2$`IBM`)^3

library(psych) # Need to use skew() in psych package
skew(dt2$`IBM`)
```

# Kurtosis
```{r}
# 3 is deducted from the eq. because kurtosis for a z-dist is 3
(mean((dt2$`IBM`-mean(dt2$`IBM`))^4)/sd(dt2$`IBM`)^4)-3

kurtosi(dt2$`IBM`)
```

# Descriptive statistics
```{r}
#library(psych)
describe(dt2$`IBM`)
describe(dt2$`CSCO`)

```

# Descriptive Statistics for Grouped Data
```{r}
dt %>%
  rename(size = total_price_excluding_optional_support) %>%
  group_by(poverty_level) %>%
  summarise(mean(size), median(size), var(size), sd(size), max(size), min(size))
```

```{r}
library(psych)
describeBy(dt$total_price_excluding_optional_support, group=dt$poverty_level)
```
# proprotion is commonly used to deal with categorical data
```{r}
dt %>%
  group_by(poverty_level) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n))
```

# Boxplot
```{r}
boxplot(dt2$`IBM`)
```
```{r}
boxplot(students_reached~poverty_level,data=dt)
```

# Covariance
```{r}
dt %>%
  filter(total_price_including_optional_support < 3000) %>%
  ggplot(aes(total_price_including_optional_support, total_donations)) +
  geom_point(size=0.5) +
  theme_classic() +
  xlab("Requested") +
  ylab("Donated")
```

# Use: an optional character string giving a method for computing covariances in the presence of missing values. This must be (an abbreviation of) one of the strings "everything": NAs will propagate conceptually, 
#"all.obs" = the presence of missing observations will produce an error, 
#"complete.obs"= missing values are handled by casewise deletion, 
#"na.or.complete" = the same unless there are no complete cases, or 
#"pairwise.complete.obs" = the correlation or covariance between each pair of variables is computed using all complete pairs of observations on those variables.

```{r}
cov(dt$total_price_including_optional_support,dt$total_donations, use="complete.obs")
```


# Correlation
# method: a character string indicating which correlation coefficient (or covariance) is to be computed. One of "pearson" (default), "kendall", or "": can be abbreviated.
```{r}
cor(dt$total_price_including_optional_support, dt$total_donations, use="complete.obs", method = "pearson")
```

```{r}
cor(dt$total_price_including_optional_support, dt$total_donations, use="complete.obs", method = "spearman")
```

```{r}
cor(dt$total_price_including_optional_support, dt$total_donations, use="complete.obs", method = "kendall")
```

```{r}
#library(psych)
corr.test(dt %>% select(students_reached, total_donations, num_donors, total_limited_base))
```

# Save correlation matrix by using pander package
```{r}
# install.packages("pander")
library("pander")

# save your date in a file
mycorrelation <- psych::corr.test(dt %>% select(students_reached, total_donations, num_donors))

#Prints an R object in Pandoc's markdown
mycorrelation$r %>%
  pander(caption="The table of the correlation matrix.")

write.csv(mycorrelation$r, file="mycorrelation-r-values.csv")
```

```{r}
mycorrelation
```



```{r}
library(GGally)
```
```{r}
dt %>% 
  select(students_reached, total_donations, num_donors, total_limited_base) %>% 
# A matrix of scatterplots is produced.
  pairs
```

```{r}
corPlot(dt %>% select(students_reached, total_donations, num_donors, total_limited_base))
```
```{r}
# Factors are used to represent categorical data. Factors can be ordered or unordered and are an important class for statistical analysis and for plotting.
dt %>% 
  select(students_reached, total_donations, num_donors, total_limited_base) %>%
  GGally::ggpairs()

```
```{r}
library("corrgram")

dt %>% 
  select(students_reached, total_donations, num_donors, total_limited_base) %>%
 corrgram(lower.panel=corrgram::panel.ellipse,
         upper.panel=panel.cor,
         diag.panel=panel.density)
```
# Correlation Heatmap
```{r}
ggcorr(dt %>% select(students_reached, total_donations, num_donors, total_limited_base), digits = 12, label = TRUE, name = "Correlations", label_alpha = TRUE)
```


# Explore the outliers

```{r}
dt %>%
  sample_n(100) %>%
  ggplot(aes(total_price_excluding_optional_support, total_donations)) +
  geom_point() +
  theme_classic()
```


```{r}
dt %>%
  sample_n(100) %>%
  select(total_price_excluding_optional_support) %>%
  mutate(z_score = (total_price_excluding_optional_support - mean(total_price_excluding_optional_support)) / sd(total_price_excluding_optional_support)) %>%
  filter(z_score > 3)
```







